---
title: "Assignment 4 - Coordinating Heart Rate"
author: "Riccardo Fusaroli"
date: "November 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysing Heart Rate and Respiration data

The goal of this assignment is to first familiarize you with heart rate, and respiration data and their preprocessing. The second part explores how to analyze interpersonal coordination of these signals.

These are the questions you need to be able to answer at the end of the assignment (aka that you need to submit as part of the portfolio)

1) How do you preprocess heart rate and respiration data? Describe the process. If any data needs to be excluded, list the excluded data and motivate the exclusion.

2) Do you observe interpersonal coordination in heart rate and respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.

3) Do you observe differences in coordination between conditions? Report the models and results.

4) Is respiration coordination a likely driver of heart rate coordination? Describe how you would test for it. Bonus points if you actually run the tests and report methods and results.

N.B. to give you a bit more data I included data from last year (Study1) and from your class (Study2). Note that synchronouns and turn-taking are the same across both studies, but the third condition is different: last year it was self-paced joint reading; this year it was the tv-series conversation. So you might want to exclude the self-paced reading (but, up to you!)

## Step by step suggestions to solve the assignment

### Exploring physiological signals

- Choose one pair (one pair, three conditions)
- Load the logs
- Produce a plot of the participants' respiration signal and a different one of the participants' HR signal (for inspecting whether the data is usable)
  N.B: remember the slides: artifacts, downsampling, scaling.
  N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3, ncol=3)
- Can you eye-ball which condition if any displays more physiological coordination?
```{r}
# libraries 

library(pacman)

p_load(groupdata2,gridExtra,pastecs,tidyverse,readr,crqa,dplyr,ggplot2,lmerTest,MuMIn)

# Pair 9. Three conditions. Read in files. 
setwd("C:/Users/Bruger/Google Drev/Cognitive Science/Experimental methods 3/assignment 4/data")
G9_T1 <- read.delim("Study2_G9_T1_Synchronous.csv", sep=",") 
G9_T2 <- read.delim("Study2_G9_T2_Conversation.csv", sep=",")
G9_T3 <- read.delim("Study2_G9_T3_TurnTaking.csv", sep=",")


# Step 1: Downsampling. Calculate mean for every 100 data-points.
# SYNCHRONOUS
G9_T1 = G9_T1 %>%    
  group(n= 100, method= 'greedy') %>%    
  dplyr::summarise(time= mean(time,na.rm=T), HR1 = mean(HR1,na.rm=T), HR2 = mean(HR2,na.rm=T), 
                   Resp1 = mean(Resp1,na.rm=T), Resp2 = mean(Resp2,na.rm=T))

# CONVERSATION
G9_T2 = G9_T2 %>%    
  group(n= 100, method= 'greedy') %>%    
  dplyr::summarise(time= mean(time,na.rm=T), HR1 = mean(HR1,na.rm=T), HR2 = mean(HR2,na.rm=T), 
                   Resp1 = mean(Resp1,na.rm=T), Resp2 = mean(Resp2,na.rm=T))

# TURN-TAKING
G9_T3 = G9_T3 %>%    
  group(n= 100, method= 'greedy') %>%    
  dplyr::summarise(time= mean(time,na.rm=T), HR1 = mean(HR1,na.rm=T), HR2 = mean(HR2,na.rm=T), 
                   Resp1 = mean(Resp1,na.rm=T), Resp2 = mean(Resp2,na.rm=T))

# Step 2: Outlier removal / Artifacts. A function to remove. Explain? 
removeOuts <- function(ts,threshold) {
  ts[ts > (mean(ts,na.rm=T) + 
             (threshold*sd(ts,na.rm=T))) |
       ts < (mean(ts,na.rm=T) - 
             (threshold*sd(ts,na.rm=T)))] = 
    mean(ts,na.rm=T) 
  return(ts)} 
threshold=2.5 

# Using function on HR and respiratory signals 
# SYNCHRONOUS
G9_T1$HR1 <- removeOuts(G9_T1$HR1,threshold)
G9_T1$HR2 <- removeOuts(G9_T1$HR2,threshold)
G9_T1$Resp1 <- removeOuts(G9_T1$Resp1,threshold) 
G9_T1$Resp2 <- removeOuts(G9_T1$Resp2,threshold) 

# CONVERSATION
G9_T2$HR1 <- removeOuts(G9_T2$HR1,threshold)
G9_T2$HR2 <- removeOuts(G9_T2$HR2,threshold)
G9_T2$Resp1 <- removeOuts(G9_T2$Resp1,threshold) 
G9_T2$Resp2 <- removeOuts(G9_T2$Resp2,threshold) 

# TURN-TAKING
G9_T3$HR1 <- removeOuts(G9_T3$HR1,threshold)
G9_T3$HR2 <- removeOuts(G9_T3$HR2,threshold)
G9_T3$Resp1 <- removeOuts(G9_T3$Resp1,threshold) 
G9_T3$Resp2 <- removeOuts(G9_T3$Resp2,threshold) 

# Step 3: Scale using r-function
# SYNCHRONOUS 
G9_T1$Resp1S=scale(G9_T1$Resp1)  
G9_T1$Resp2S=scale(G9_T1$Resp2)  
G9_T1$HR1S=scale(G9_T1$HR1) 
G9_T1$HR2S=scale(G9_T1$HR2) 

# CONVERSATION
G9_T2$Resp1S=scale(G9_T2$Resp1)  
G9_T2$Resp2S=scale(G9_T2$Resp2)  
G9_T2$HR1S=scale(G9_T2$HR1) 
G9_T2$HR2S=scale(G9_T2$HR2) 

# TURN-TAKING
G9_T3$Resp1S=scale(G9_T3$Resp1)  
G9_T3$Resp2S=scale(G9_T3$Resp2)  
G9_T3$HR1S=scale(G9_T3$HR1) 
G9_T3$HR2S=scale(G9_T3$HR2) 

# plot HR signal (both in one plot) with the preprocessed data
# SYNCRONOUS
plot_sync_HR <- ggplot(G9_T1, aes(x=time, y=HR1S),colour="darkgreen")+
  geom_line()+
  geom_line(aes(x=time, y=HR2S),colour = "darkred")+
  labs(x="time",y="HR",header="HR Synchronous: Pair 9")

# CONVERSATION
plot_conv_HR <- ggplot(G9_T2, aes(x=time, y=HR1S),colour="darkgreen")+
  geom_line()+
  geom_line(aes(x=time, y=HR2S),colour = "darkred")+
  labs(x="time",y="HR",header="HR Conversation: Pair 9")

# TURN-TAKING
plot_turn_HR <- ggplot(G9_T3, aes(x=time, y=HR1S),colour="darkgreen")+
  geom_line()+
  geom_line(aes(x=time, y=HR2S),colour = "darkred")+
  labs(x="time",y="HR",header="HR Turn-taking: Pair 9")


# plot respiratory signal (both in one plot) with the preprocessed data
# SYNCHRONOUS
plot_sync_Resp <- ggplot(G9_T1, aes(x=time, y=Resp1S),colour="darkgreen")+
  geom_line()+
  geom_line(aes(x=time, y=Resp2S),colour = "darkred")+
  labs(x="time",y="Respiration",header="Respiration signal Synchronous: Pair 9")

# CONVERSATION
plot_conv_Resp <- ggplot(G9_T2, aes(x=time, y=Resp1S),colour="darkgreen")+
  geom_line()+
  geom_line(aes(x=time, y=Resp2S),colour = "darkred")+
  labs(x="time",y="Respiration",header="Respiration signal Conversation: Pair 9")

# TURN-TAKING
plot_turn_Resp <- ggplot(G9_T3, aes(x=time, y=Resp1S),colour="darkgreen")+
  geom_line()+
  geom_line(aes(x=time, y=Resp2S),colour = "darkred")+
  labs(x="time",y="Respiration",header="Respiration signal Turn-taking: Pair 9")

# Arrangeing plots HR
gridExtra::grid.arrange(plot_sync_HR,plot_conv_HR,plot_turn_HR, ncol=3)

# Arrangeing plots Respiration
gridExtra::grid.arrange(plot_sync_Resp,plot_conv_Resp,plot_turn_Resp, ncol=3)

# Eye-ball which condition if any displays more physiological coordination.
# Well pretty hard to see anything.....

```

- Run crqa on heart rate and respiration data (find parameters, run crqa)
- Does this tell you more than just eyeballing the plots?
```{r}


par <- list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0, mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") 


# SYNCHRONOUS
A_sync <- optimizeParam(G9_T1$HR1S, G9_T1$HR2S, par, min.rec= 2, max.rec= 8)
B_sync <- optimizeParam(G9_T1$Resp1S, G9_T1$Resp2S, par, min.rec= 2, max.rec= 8)

# d <- 
# e <- 
# r <- 

#C <- crqa(G9_T1$HR1S, G9_T1$HR2S, delay=d, embed=e, radius=r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
#D <- crqa(G9_T1$Resp1S, G9_T1$Resp2S, delay=d, embed=e, radius=r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)

# CONVERSATION
A_conv <- optimizeParam(G9_T2$HR1S, G9_T2$HR2S, par, min.rec= 2, max.rec= 8)
B_conv <- optimizeParam(G9_T2$Resp1S, G9_T2$Resp2S, par, min.rec= 2, max.rec= 8)

# d <- 
# e <- 
# r <- 

#C <- crqa(G9_T1$HR1S, G9_T1$HR2S, delay=d, embed=e, radius=r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
#D <- crqa(G9_T1$Resp1, G9_T1$Resp2, delay=d, embed=e, radius=r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)

# TURNTAKING
A_turn <- optimizeParam(G9_T3$HR1S, G9_T3$HR2S, par, min.rec= 2, max.rec= 8)
B_turn <- optimizeParam(G9_T3$Resp1S, G9_T3$Resp2S, par, min.rec= 2, max.rec= 8)

# d <- 
# e <- 
# r <- 

#C <- crqa(G9_T1$HR1S, G9_T1$HR2S, delay=d, embed=e, radius=r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
#D <- crqa(G9_T1$Resp1, G9_T1$Resp2, delay=d, embed=e, radius=r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)


```


### Systematically pre-process the data
- Loop through all the files (either with a loop or with a function), check which files should be excluded, if any, and save the pre-processed time-series. Tip: plot and visually inspect the data to figure out which should be excluded.

```{r}

setwd("C:/Users/Bruger/Google Drev/Cognitive Science/Experimental methods 3/assignment 4")


file_list <- list.files(path = "data", full.names = T)

  n=1
  for (i in 1:length(file_list)) {
  
    print(i)

  # read in file 
    file <- read.delim(file_list[i],sep = ",") 
    
  
  # Down sample
    file = file %>%    
    group(n= 100, method= 'greedy') %>%    
    dplyr::summarise(time= mean(time,na.rm=T), HR1 = mean(HR1,na.rm=T), HR2 = mean(HR2,na.rm=T), 
                   Resp1 = mean(Resp1,na.rm=T), Resp2 = mean(Resp2,na.rm=T))
  
  # Artefacts. remove outliers 
    removeOuts <- function(ts,threshold) {
    ts[ts > (mean(ts,na.rm=T) + 
             (threshold*sd(ts,na.rm=T))) |
       ts < (mean(ts,na.rm=T) - 
             (threshold*sd(ts,na.rm=T)))] = 
    mean(ts,na.rm=T) 
    return(ts)} 
    threshold=2.5 

    file$HR1 <- removeOuts(file$HR1,threshold)
    file$HR2 <- removeOuts(file$HR2,threshold)
    file$Resp1 <- removeOuts(file$Resp1,threshold) 
    file$Resp2 <- removeOuts(file$Resp2,threshold) 
  
  # Scale 
    file$HR1S=scale(file$HR1) 
    file$HR2S=scale(file$HR2)
    file$Resp1S=scale(file$Resp1)  
    file$Resp2S=scale(file$Resp2)  
  
    # name file 
    file1 = paste0("pp_", file_list[i])
    # save df
    write.csv(file, file.path(file1))
    
    
  n=n+1
  
}
  

```

```{r}

# loop and plot

pp_file_list <- list.files(path = "pp_data", full.names = T)

n=1
for (i in 1:length(pp_file_list)){
  
  # read in file 
  p <- read.delim(pp_file_list[i],sep = ",")
  
  #name file
  name_HR = paste0("HR_", n, ".pdf")
  
  # plot HR
  HR_plot <- ggplot(p, aes(x=time, y=HR1S),colour="lightblue")+
  geom_line()+
  geom_line(aes(x=time, y=HR2S),colour = "red")+
  labs(x="time",y="HR",header=file.path(name_HR))
  
  # save HR plot
  ggsave(filename = file.path(name_HR), device = "pdf")
  
  #name file
  name_resp = paste0("Resp_",n,".pdf")
  
  # plot Resp
  ggplot(p, aes(x=time, y=Resp1S),colour="green")+
  geom_line()+
  geom_line(aes(x=time, y=Resp2S),colour = "pink")+
  labs(x="time",y="Respiration",header=file.path(name_resp))
  
  # save Resp plot
  ggsave(filename = file.path(name_resp), device = "pdf")

  n=n+1  
}


# exclude resp 1 and 2 - Study 1 G1 Synch and Turn. + G2
# -> remove all Study 1 G1 and G2




```


- Run crqa on all the pre-processed time-series and save the output (don't forget to add columns with study, group, condition and trial). Tip: remember to first assess optimal parameters (dimensions, delay, radius) across all timeseries. Tip: it will often fail, just take whatever parameters you get, select optimal across timeseries parameters and run crqa on all timeseries with those. Tip: double check the rr. When I ran the loop, I got very low rr, so I adjusted the radius until the average of rr across all pairs was approx. 4%.

```{r}


#Loop 1 going through all the files and trying to calculate optimal delay, embedding and radius for HR and resp separately. Save them in a dataframe.

HR_delay = NULL
HR_embed = NULL
HR_radius = NULL 

Resp_delay = NULL
Resp_embed = NULL
Resp_radius = NULL

pp_file_list <- list.files(path = "pp_data", full.names = T)

test_1 <- read.delim(pp_file_list[1],sep = ",")

n = 1
for (i in 1:length(pp_file_list)){
  
  # read in file 
  f <- read.delim(pp_file_list[i],sep = ",")
  
  # Parameters 
  par <- list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0, mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") 
  
  # Optimise parameters HR  
  A <- try(optimizeParam(f$HR1S, f$HR2S, par, min.rec= 2, max.rec= 8))
  # Optimise parameters Resp
  B <- try(optimizeParam(f$Resp1S, f$Resp2S, par, min.rec= 2, max.rec= 8))
  
  # Put parameters HR into lists 
  if(length(A)<2){
    HR_delay[n]=NA
    HR_embed[n]=NA
    HR_radius[n]=NA
    }
  else{
    HR_delay[n]=A$delay[1]
    HR_embed[n]=A$emddim[1]
    HR_radius[n]=A$radius[1]
  }  
  
  # Put parameters Resp into lists 
  if(length(B)<2){
    Resp_delay[n]=NA
    Resp_embed[n]=NA
    Resp_radius[n]=NA
    }
  else{
    Resp_delay[n]=B$delay[1]
    Resp_embed[n]=B$emddim[1]
    Resp_radius[n]=B $radius[1]
    } 

n=n+1   
}

df_parameters <- data.frame(HR_delay,HR_embed, HR_radius, Resp_delay, Resp_embed, Resp_radius)


    
```

```{r}

# Choose 1 delay for HR (all files), 1 embedding, 1 radius; same for resp

HR_d <- median(df_parameters$HR_delay,na.rm = T)
HR_e <- median(df_parameters$HR_embed,na.rm = T)
HR_r <- median(df_parameters$HR_radius,na.rm = T)+0.8 

Resp_d <- median(df_parameters$Resp_delay,na.rm = T)
Resp_e <- median(df_parameters$Resp_embed,na.rm = T)
Resp_r <- median(df_parameters$Resp_radius,na.rm = T)

# Loop 2 use the chosen values to run crqa on all files (same parameters for all files). Save crqa results. Pair, condition, trial, HRrr, HRdet,.., Resprr

group = NULL
study = NULL
trial = NULL
condition = NULL

HR_RR = NULL
HR_DET = NULL
HR_maxL= NULL
HR_L = NULL
HR_ENTR = NULL
HR_LAM = NULL
HR_TT = NULL

Resp_RR = NULL
Resp_DET = NULL
Resp_maxL= NULL
Resp_L = NULL
Resp_ENTR = NULL
Resp_LAM = NULL
Resp_TT = NULL

pp_file_list <- list.files(path = "pp_data", full.names = T)

n = 1
for (i in 1:length(pp_file_list)){
  
  # read in file 
  d <- read.delim(pp_file_list[i],sep = ",")
  
  # Extract descriptors
  group[n] = str_extract(pp_file_list[i],regex("G\\d{1,2}"))
  study[n] = str_extract(pp_file_list[i], regex("Study\\d{1}"))
  trial[n] = str_extract(pp_file_list[i],regex("T\\d{1}"))
  condition[n] = str_extract(pp_file_list[i],regex("[A-Z a-z]{6,}"))
  
  # Run crqa HR
  C <- try(crqa(d$HR1S, d$HR2S, delay=HR_d, embed=HR_e, radius=HR_r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  if(length(C)<2){
    HR_RR[n] = NA
    HR_DET[n] = NA
    HR_maxL[n] = NA
    HR_L[n] = NA
    HR_ENTR[n] = NA
    HR_LAM[n] = NA
    HR_TT[n] = NA
  }
  
  else{
    HR_RR[n] = C$RR[1]
    HR_DET[n] = C$DET[1]
    HR_maxL[n] = C$maxL[1]
    HR_L[n] = C$L[1]
    HR_ENTR[n] = C$ENTR[1]
    HR_LAM[n] = C$LAM[1]
    HR_TT[n] = C$TT[1]
  }
  
    # Run crqa Resp
  E <- try(crqa(d$Resp1S, d$Resp2S, delay=Resp_d, embed=Resp_e, radius=Resp_r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  if(length(E)<2){
    Resp_RR[n] = NA
    Resp_DET[n] = NA
    Resp_maxL[n] = NA
    Resp_L[n] = NA
    Resp_ENTR[n] = NA
    Resp_LAM[n] = NA
    Resp_TT[n] = NA
  }
  
  else{
    Resp_RR[n] = E$RR[1]
    Resp_DET[n] = E$DET[1]
    Resp_maxL[n] = E$maxL[1]
    Resp_L[n] = E$L[1]
    Resp_ENTR[n] = E$ENTR[1]
    Resp_LAM[n] = E$LAM[1]
    Resp_TT[n] = E$TT[1]
    }
  
  n=n+1
}

rqa_real <- data.frame(group, study, trial, condition, HR_RR, HR_DET, HR_maxL, HR_L, HR_ENTR, HR_LAM, HR_TT, Resp_RR, Resp_DET, Resp_maxL, Resp_L, Resp_ENTR, Resp_LAM, Resp_TT)

write.csv(rqa_real,file="rqa_real.csv")


```


### Creating controls: shuffled controls
 - loop through all pairs and conditions
 - shuffle the timeseries (take a timeseries and rearrange its values in a random order). Tip check the sample() function
 - run crqa and save the output. NB. which delay, embed, radius parameters should you use?
 # the ones from before? for easy comparison?
 - statistically compare the crqa indexes in real and shuffled pairs
 
```{r}

# shuffle rows in each df and save as shuffled data 
pp_file_list <- list.files(path = "pp_data", full.names = T)

n = 1
for (i in 1:length(pp_file_list)){
  
  # read in file 
  f <- read.delim(pp_file_list[i],sep = ",")
  
  f$HR1SH <- sample(f$HR1S,nrow(f)) 
  f$HR2SH <- sample(f$HR2S,nrow(f)) 
  f$Resp1SH <- sample(f$Resp1S,nrow(f)) 
  f$Resp2SH <- sample(f$Resp2S,nrow(f)) 
   
  # name file 
  file1 = paste0("shuffled_", pp_file_list[i])
  # save df
  write.csv(f, file.path(file1))
  
  
  n=n+1 
}


```
 
```{r}

group = NULL
study = NULL
trial = NULL
condition = NULL

shuff_HR_RR = NULL
shuff_HR_DET = NULL
shuff_HR_maxL= NULL
shuff_HR_L = NULL
shuff_HR_ENTR = NULL
shuff_HR_LAM = NULL
shuff_HR_TT = NULL

shuff_Resp_RR = NULL
shuff_Resp_DET = NULL
shuff_Resp_maxL= NULL
shuff_Resp_L = NULL
shuff_Resp_ENTR = NULL
shuff_Resp_LAM = NULL
shuff_Resp_TT = NULL

shuff_file_list <- list.files(path = "shuffled_pp_data", full.names = T)

n = 1
for (i in 1:length(shuff_file_list)){
  
  # read in file 
  d <- read.delim(shuff_file_list[i],sep = ",")
  
  # Extract descriptors
  group[n] = str_extract(shuff_file_list[i],regex("G\\d{1,2}"))
  study[n] = str_extract(shuff_file_list[i], regex("Study\\d{1}"))
  trial[n] = str_extract(shuff_file_list[i],regex("T\\d{1}"))
  condition[n] = str_extract(shuff_file_list[i],regex("[A-Z a-z]{9,}"))
  
  # Run crqa HR
  G <- try(crqa(d$HR1SH, d$HR2SH, delay=HR_d, embed=HR_e, radius=HR_r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  if(length(G)<2){
    shuff_HR_RR[n] = NA
    shuff_HR_DET[n] = NA
    shuff_HR_maxL[n] = NA
    shuff_HR_L[n] = NA
    shuff_HR_ENTR[n] = NA
    shuff_HR_LAM[n] = NA
    shuff_HR_TT[n] = NA
  }
  
  else{
    shuff_HR_RR[n] = G$RR[1]
    shuff_HR_DET[n] = G$DET[1]
    shuff_HR_maxL[n] = G$maxL[1]
    shuff_HR_L[n] = G$L[1]
    shuff_HR_ENTR[n] = G$ENTR[1]
    shuff_HR_LAM[n] = G$LAM[1]
    shuff_HR_TT[n] = G$TT[1]
  }
  
    # Run crqa Resp
  H <- try(crqa(d$Resp1SH, d$Resp2SH, delay=Resp_d, embed=Resp_e, radius=Resp_r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  if(length(H)<2){
    shuff_Resp_RR[n] = NA
    shuff_Resp_DET[n] = NA
    shuff_Resp_maxL[n] = NA
    shuff_Resp_L[n] = NA
    shuff_Resp_ENTR[n] = NA
    shuff_Resp_LAM[n] = NA
    shuff_Resp_TT[n] = NA
  }
  
  else{
    shuff_Resp_RR[n] = H$RR[1]
    shuff_Resp_DET[n] = H$DET[1]
    shuff_Resp_maxL[n] = H$maxL[1]
    shuff_Resp_L[n] = H$L[1]
    shuff_Resp_ENTR[n] = H$ENTR[1]
    shuff_Resp_LAM[n] = H$LAM[1]
    shuff_Resp_TT[n] = H$TT[1]
    }
  
  n=n+1
}

rqa_shuffled <- data.frame(group, study, trial, condition, shuff_HR_RR, shuff_HR_DET, shuff_HR_maxL, shuff_HR_L, shuff_HR_ENTR, shuff_HR_LAM, shuff_HR_TT, shuff_Resp_RR, shuff_Resp_DET, shuff_Resp_maxL, shuff_Resp_L, shuff_Resp_ENTR, shuff_Resp_LAM, shuff_Resp_TT)

write.csv(rqa_shuffled,file="rqa_shuffled.csv")

```
 
 
```{r}

# compare rqa_real and rqa_shuffled

```
 
### TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair). Tip: Celine will share a commented script
 - Run crqa on all the surrogate pairs and save the output. NB. which delay, embed, radius parameters should you use?
 - Test whether crqa shows a difference between real and surrogate pairs
 
```{r}

setwd("C:/Users/Bruger/Google Drev/Cognitive Science/Experimental methods 3/assignment 4")
prepro_filelist = list.files(pattern = "Study1") # load the preprocessed files
 
num=str_extract_all(prepro_filelist, "\\d+") #extract all numbers in the file name (i.e Study?_G?_T?)

df= as.data.frame(t(as.data.frame(num))) #create a transposed (i.e. "tipped") dataframe of these number
rownames(df)<-NULL #Remove rownames.

pairs=as.numeric(unique(as.numeric(as.character(df$V2)))) #create pairs, i.e. unique group numbers

SurrogateList <- expand.grid(a = pairs, b = pairs) # expand.grid combines all possible combinations of a given vector
SurrogateList = subset(SurrogateList, a != b) # Ensure that pair 1 (i.e group1) doesn't get paired with it self.


# Per each condition
for (co in c("Synchronous","TurnTaking")){ #loop through a list of condintions 
  for (i in 1:nrow(SurrogateList)){ #and loop through all the pairings of the surrogate list
    
  #start recreate filenames from the working directory based on the matching from the surrogatelist
  f1=paste0('Study1_G',SurrogateList$a[i]) #Insert group numeber from Surrogatelist 
  f2=paste0('Study1_G',SurrogateList$b[i])
  
  file1=prepro_filelist[grepl(f1,prepro_filelist) & grepl(co,prepro_filelist)] #insert conditions as well.
  file2=prepro_filelist[grepl(f2,prepro_filelist) & grepl(co,prepro_filelist)]
  
  d1=read.csv(file1) #Load the recreated filenames
  
  d2=read.csv(file2) 
  
  #Ensure both dataframes have the same length
  if (nrow(d1)>nrow(d2)){
    d1=d1[1:nrow(d2),]
  } else {
    d2=d2[1:nrow(d1),]
  }
  
  #creation of surrogate files. (Riccardo's idea of having fun and completely overdoing it ;-) ) 
  #per each 2 pairs there are 4 possible surrogate pairs: 
  #pair1person1 w pair2person2; 
  d3=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp2,HR1=d1$HR1,HR2=d2$HR2,Resp1S=d1$Resp1S,Resp2S=d2$Resp2S,HR1S=d1$HR1S,HR2S=d2$HR2S)
  #pair1person2 w pair2person2;
  d4=data.frame(Resp1=d2$Resp1,Resp2=d1$Resp2,HR1=d2$HR1,HR2=d1$HR2,Resp1S=d2$Resp1S,Resp2S=d1$Resp2S,HR1S=d2$HR1S,HR2S=d1$HR2S)
  
  #pair1person1 w pair2person1;
  d5=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp1,HR1=d1$HR1,HR2=d2$HR1,Resp1S=d1$Resp1S,Resp2S=d2$Resp1S,HR1S=d1$HR1S,HR2S=d2$HR1S)
  #pair1person2 w pair2person2
  d6=data.frame(Resp1=d2$Resp2,Resp2=d1$Resp2,HR1=d2$HR2,HR2=d1$HR2,Resp1S=d2$Resp2S,Resp2S=d1$Resp2S,HR1S=d2$HR2S,HR2S=d1$HR2S)
  
  #Make filenames that tells which pairs have been mixed
  file3=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"a_",co,".csv")
  file4=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"b_",co,".csv")
  file5=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"c_",co,".csv")
  file6=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"d_",co,".csv")
  
  #write csv files
  write_csv(d3,file.path("surrogate",file3))
  write_csv(d4,file.path("surrogate",file4))
  write_csv(d5,file.path("surrogate",file5))
  write_csv(d6,file.path("surrogate",file6))
  }
}



```

```{r}
##NOW DO THE SAME FOR STUDY 2##

setwd("C:/Users/Bruger/Google Drev/Cognitive Science/Experimental methods 3/assignment 4")
prepro_filelist = list.files(pattern = "Study2") # load the preprocessed files

num=str_extract_all(prepro_filelist, "\\d+") #extract all numbers in the file name (i.e Study?_G?_T?)

df= as.data.frame(t(as.data.frame(num))) #create a transposed (i.e. "tipped") dataframe of these number
rownames(df)<-NULL #Remove rownames.

pairs=as.numeric(unique(as.numeric(as.character(df$V2)))) #create pairs, i.e. unique group numbers

SurrogateList <- expand.grid(a = pairs, b = pairs) # expand.grid combines all possible combinations of a given vector
SurrogateList = subset(SurrogateList, a != b) # Ensure that pair 1 (i.e group1) doesn't get paired with it self.


# Per each condition
for (co in c("Synchronous","TurnTaking","Conversation")){ #loop through a list of condintions 
  for (i in 1:nrow(SurrogateList)){ #and loop through all the pairings of the surrogate list
    
  #start recreate filenames from the working directory based on the matching from the surrogatelist
  f1=paste0('Study2_G',SurrogateList$a[i]) #Insert group numeber from Surrogatelist 
  f2=paste0('Study2_G',SurrogateList$b[i])
  
  file1=prepro_filelist[grepl(f1,prepro_filelist) & grepl(co,prepro_filelist)] #insert conditions as well.
  file2=prepro_filelist[grepl(f2,prepro_filelist) & grepl(co,prepro_filelist)]
  
  d1=read.csv(file1) #Load the recreated filenames
  
  d2=read.csv(file2) 
  
  #Ensure both dataframes have the same length
  if (nrow(d1)>nrow(d2)){
    d1=d1[1:nrow(d2),]
  } else {
    d2=d2[1:nrow(d1),]
  }
  
  #creation of surrogate files. (Riccardo's idea of having fun and completely overdoing it ;-) ) 
  #per each 2 pairs there are 4 possible surrogate pairs: 
  #pair1person1 w pair2person2; 
  d3=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp2,HR1=d1$HR1,HR2=d2$HR2,Resp1S=d1$Resp1S,Resp2S=d2$Resp2S,HR1S=d1$HR1S,HR2S=d2$HR2S)
  #pair1person2 w pair2person2;
  d4=data.frame(Resp1=d2$Resp1,Resp2=d1$Resp2,HR1=d2$HR1,HR2=d1$HR2,Resp1S=d2$Resp1S,Resp2S=d1$Resp2S,HR1S=d2$HR1S,HR2S=d1$HR2S)
  
  #pair1person1 w pair2person1;
  d5=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp1,HR1=d1$HR1,HR2=d2$HR1,Resp1S=d1$Resp1S,Resp2S=d2$Resp1S,HR1S=d1$HR1S,HR2S=d2$HR1S)
  #pair1person2 w pair2person2
  d6=data.frame(Resp1=d2$Resp2,Resp2=d1$Resp2,HR1=d2$HR2,HR2=d1$HR2,Resp1S=d2$Resp2S,Resp2S=d1$Resp2S,HR1S=d2$HR2S,HR2S=d1$HR2S)
  
  #Make filenames that tells which pairs have been mixed
  file3=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"a_",co,".csv")
  file4=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"b_",co,".csv")
  file5=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"c_",co,".csv")
  file6=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"d_",co,".csv")
  
  #write csv files
  write_csv(d3,file.path("surrogate",file3))
  write_csv(d4,file.path("surrogate",file4))
  write_csv(d5,file.path("surrogate",file5))
  write_csv(d6,file.path("surrogate",file6))
  }
}

```

```{r}
# rqa surrogate 

group = NULL
study = NULL
trial = NULL
condition = NULL

sur_HR_RR = NULL
sur_HR_DET = NULL
sur_HR_maxL= NULL
sur_HR_L = NULL
sur_HR_ENTR = NULL
sur_HR_LAM = NULL
sur_HR_TT = NULL

sur_Resp_RR = NULL
sur_Resp_DET = NULL
sur_Resp_maxL= NULL
sur_Resp_L = NULL
sur_Resp_ENTR = NULL
sur_Resp_LAM = NULL
sur_Resp_TT = NULL

sur_file_list <- list.files(path = "surrogate", full.names = T)

test <- read.delim(sur_file_list[1],sep = ",")

n = 1
for (i in 1:length(sur_file_list)){
  
  # read in file 
  d <- read.delim(sur_file_list[i],sep = ",")
  
  # Extract descriptors
  group[n] = str_extract(sur_file_list[i],regex("\\d{1,2}&\\d{1,2}[a-z]{1}"))
  study[n] = str_extract(sur_file_list[i], regex("Study\\d{1}"))
  trial[n] = NA
  condition[n] = str_extract(sur_file_list[i],regex("Synchronous|TurnTaking|Conversation"))
  
  # Run crqa HR
  J <- try(crqa(d$HR1S, d$HR2S, delay=HR_d, embed=HR_e, radius=HR_r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  if(length(J)<2){
    sur_HR_RR[n] = NA
    sur_HR_DET[n] = NA
    sur_HR_maxL[n] = NA
    sur_HR_L[n] = NA
    sur_HR_ENTR[n] = NA
    sur_HR_LAM[n] = NA
    sur_HR_TT[n] = NA
  }
  
  else{
    sur_HR_RR[n] = J$RR[1]
    sur_HR_DET[n] = J$DET[1]
    sur_HR_maxL[n] = J$maxL[1]
    sur_HR_L[n] = J$L[1]
    sur_HR_ENTR[n] = J$ENTR[1]
    sur_HR_LAM[n] = J$LAM[1]
    sur_HR_TT[n] = J$TT[1]
  }
  
    # Run crqa Resp
  K <- try(crqa(d$Resp1S, d$Resp2S, delay=Resp_d, embed=Resp_e, radius=Resp_r,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  if(length(K)<2){
    sur_Resp_RR[n] = NA
    sur_Resp_DET[n] = NA
    sur_Resp_maxL[n] = NA
    sur_Resp_L[n] = NA
    sur_Resp_ENTR[n] = NA
    sur_Resp_LAM[n] = NA
    sur_Resp_TT[n] = NA
  }
  
  else{
    sur_Resp_RR[n] = K$RR[1]
    sur_Resp_DET[n] = K$DET[1]
    sur_Resp_maxL[n] = K$maxL[1]
    sur_Resp_L[n] = K$L[1]
    sur_Resp_ENTR[n] = K$ENTR[1]
    sur_Resp_LAM[n] = K$LAM[1]
    sur_Resp_TT[n] = K$TT[1]
    }
  
  n=n+1
}

rqa_sur <- data.frame(group, study, trial, condition, sur_HR_RR, sur_HR_DET, sur_HR_maxL, sur_HR_L, sur_HR_ENTR, sur_HR_LAM, sur_HR_TT, sur_Resp_RR, sur_Resp_DET, sur_Resp_maxL, sur_Resp_L, sur_Resp_ENTR, sur_Resp_LAM, sur_Resp_TT)

write.csv(rqa_sur,file="rqa_sur.csv")

```

```{r}
# Clean data and make one nice and pretty data frame 

rqa_real <- read.csv("rqa_real.csv")
rqa_shuffled <- read.csv("rqa_shuffled.csv")
rqa_sur <- read.csv("rqa_sur.csv")

type = "real"
rqa_real <- cbind(rqa_real,type)
rqa_real
type = "shuffled"
rqa_shuffled = cbind(rqa_shuffled,type)

type = "surrogate"
rqa_sur <- cbind(rqa_sur,type)

names(rqa_real) <- c("x","group","study","trial","condition","HR_RR","HR_DET","HR_maxL","HR_L","HR_ENTR","HR_LAM","HR_TT","Resp_RR","Resp_DET","Resp_maxL","Resp_L","Resp_ENTR","Resp_LAM","Resp_TT","type")

names(rqa_shuffled) <- c("x","group","study","trial","condition","HR_RR","HR_DET","HR_maxL","HR_L","HR_ENTR","HR_LAM","HR_TT","Resp_RR","Resp_DET","Resp_maxL","Resp_L","Resp_ENTR","Resp_LAM","Resp_TT","type")

names(rqa_sur) <- c("x","group","study","trial","condition","HR_RR","HR_DET","HR_maxL","HR_L","HR_ENTR","HR_LAM","HR_TT","Resp_RR","Resp_DET","Resp_maxL","Resp_L","Resp_ENTR","Resp_LAM","Resp_TT","type")


df_final <- rbind(rqa_real,rqa_shuffled,rqa_sur)

df_final = df_final[!df_final$condition=="SelfPaced",]

df_final$group <- as.factor(df_final$group) 

```

### Testing effects of conditions
 - make a (probably underpowered) mixed model testing effects of the different conditions on heart rate and respiration coordination
 - N.B: would it make sense to include surrogate pairs? and if so how? what would that tell you?

```{r}

# does type predict RR?
model1 = lmer(HR_RR ~ type + (1|study) + (1|group), data = df_final, REML = F)
summary(model1)
#Fixed effects:
#              Estimate Std. Error       df t value Pr(>|t|)    
#(Intercept)     5.5711     1.1705  82.1000   4.759 8.22e-06 ***
#typeshuffled   -3.0643     1.3497 376.0000  -2.270   0.0238 *  
#typesurrogate  -0.5993     1.1570 219.2000  -0.518   0.6050   

model2 = lmer(Resp_RR ~ type + (1|study) + (1|group), data = df_final, REML = F)
summary(model2)  
#(Intercept)     4.2844     0.9326  32.3000   4.594 6.33e-05 ***
#typeshuffled    0.4421     1.1327 502.4000   0.390    0.696    
#typesurrogate   0.2181     0.8528 501.9000   0.256    0.798    

# does condition predict RR?
# HR
model3 = lmer(HR_RR ~ condition + (1|study) + (1|group), subset(df_final,type=="real"), REML = F)
summary(model3)

model4 = lmer(HR_RR ~ condition + (1|study) + (1|group), subset(df_final,type=="shuffled"), REML = F)
summary(model4)

model5 = lmer(HR_RR ~ condition + (1|study) + (1|group), subset(df_final,type=="surrogate"), REML = F)
summary(model5)

# RESP
model6 = lmer(Resp_RR ~ condition + (1|study) + (1|group), subset(df_final,type=="real"), REML = F)
summary(model6)
# condition conversation is significant predicter, the others are not ?

model7 = lmer(Resp_RR ~ condition + (1|study) + (1|group), subset(df_final,type=="shuffled"), REML = F)
summary(model7)

model8 = lmer(Resp_RR ~ condition + (1|study) + (1|group), subset(df_final,type=="surrogate"), REML = F)
summary(model8)

# does the interaction between type and condition differ?
model9 = lmer(HR_RR ~ condition*type + (1|study) + (1|group), data = df_final, REML = F)
summary(model9)

model10 = lmer(Resp_RR ~ condition*type + (1|study) + (1|group), data = df_final, REML = F)
summary(model10)

# subset or interaction?
## subset: real. difference between conditions
## interaction: surrogate should have same effect as real in conversation but not synchronous?

# which crqa-measure should we predict? 
## any is fine here, but in a real project we should test more, as they measure different things 

# which outcome should we compare over types?
## subset: if any condition significantly predict RR
## interaction: we would expect sync to have a small difference between surrogate and real since we tell them how to speak/breath. and we would expect conv to have a large difference between surrogate and real since they themeself in the pair dicide speaking/breathing rythm. 

# how to statistically compare types. 
## model predict RR from type 

# should I change radius since RR is so low?
#increase radius HR by 1.2 

```


### Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them

```{r}
baseline = lmer(HR_RR ~ 1 + (1|study) + (1|group), data = df_final, REML = F)
m0 = lmer(HR_RR ~ condition + (1|study) + (1|group), data = df_final, REML = F)
summary(m0)
m1 = lmer(HR_RR ~ condition + Resp_RR + (1|study) + (1|group), data = df_final, REML = F)
summary(m1)

anova(baseline,m0,m1)



```
 
 